---
title: "Load the data"
output: html_notebook
---

```{r}
rm(list=ls())
library(readr)
library(dplyr)
library(tidyr)
library(phyloseq)
```

# Isolate and clinical data
Load the isolate and clinical data and change the layout of the data

```{r}
pats_file <- 'data/isolate_clinical_data.txt'
pat_data <- read_delim(pats_file,'\t')

# Convert percentages to proportions
tmp_data <- pat_data %>% mutate(R_AP=R_AP/100,R_FQ=R_FQ/100,R_3GC=R_3GC/100,R_AG=R_AG/100)

hospital <- tmp_data %>% filter(Source == "Hospital")
wwtp <- tmp_data %>% filter(Source == "Wastewater")

# Change names of isolate and lcinical data
hospital_resistance <- hospital %>% dplyr::select("Country","R_AP","R_FQ","R_3GC","R_AG","MAR")
colnames(hospital_resistance) <- c("Country","cl_R_AP","cl_R_FQ","cl_R_3GC","cl_R_AG","cl_R_MAR")
wwtp_resistance <- wwtp %>% dplyr::select("Country","R_AP","R_FQ","R_3GC","R_AG","MAR")
colnames(wwtp_resistance) <- c("Country","wwtp_R_AP","wwtp_R_FQ","wwtp_R_3GC","wwtp_R_AG","wwtp_R_MAR")

isolate_merged <- merge(wwtp_resistance,hospital_resistance, by = "Country")
```

# Metagenomic data

## Load the metagenomic data and extract counts and lineages
```{r}
# Paths to data files
meta_data_file <- 'data/metadata.txt'
resfinder_count_file <- 'data/semar_sample_lineage_arg_counts_90.txt'


# Load the meta data and resfinder data
meta_data <- read_delim(meta_data_file,'\t')
arg_data <- read_delim(resfinder_count_file,'\t')


# Convert the arg read counts to fragments
n_columns = dim(arg_data)[2]
n_samples <- n_columns -1
arg_counts_fragments <- ceiling(arg_data[,2:n_columns]/2)
first_column <- colnames(arg_counts_fragments)[1]
arg_data_fragments <- arg_counts_fragments %>% mutate(Gene=arg_data$Gene, .before=first_column)
```

```{r}
# Get the counts together with lineage
lineage_to_use <- c("Class","cl_90","Family", "Gene", "ARG")
arg_data_lineage <- arg_data_fragments %>% separate(Gene,lineage_to_use,';')
rownames(arg_data_lineage) <- arg_data_lineage$ARG

# Remove the disinfectant resistance genes
arg_data_lineage <- arg_data_lineage %>% filter(!Class == "Disinfectant")

# Get only the lineages
arg_lineage <- arg_data_lineage %>% dplyr::select(all_of(lineage_to_use))


# Get only the counts
arg_counts <- arg_data_lineage %>% dplyr::select(-all_of(lineage_to_use))
rownames(arg_counts) <- arg_data_lineage$ARG
```
## Normalize  the ARG count by ARG lengths

```{r}
# Normalize the args by arg length (if we want to compare different genes against each other)
ARG_lengths <- read_delim("data/gene_length.txt",'\t')

# join data frame with read lengths
arg_data_tmp <- inner_join(arg_data_lineage,ARG_lengths,by='ARG')
rownames(arg_data_tmp) <- arg_data_tmp$ARG


#extract only the counts together with the arg length
arg_counts_tmp <- arg_data_tmp %>% dplyr::select(-Class,-cl_90,-Family,-Gene,-ARG)


# Normalize each cell by the corresponding arg length
n <- ncol(arg_counts_tmp)
arg_stop <- n-1
length_normalized_args <- sweep(arg_counts_tmp[,1:arg_stop], 1,
      arg_counts_tmp$length, FUN="/")
rownames(length_normalized_args) <- arg_data_lineage$ARG

```

## Normalize the ARG count data and add it to a dataframe
```{r}
# Get total arg counts length normalized and add to meta_data frame
res_count <- apply(length_normalized_args,2,sum)

row.names(meta_data) <- meta_data$Sample
res_frame <- data.frame(res_count)
res_frame <- res_frame %>% mutate(Sample = rownames(res_frame))
new_meta_data <- merge(meta_data, res_frame, by="Sample")
rownames(new_meta_data) <- new_meta_data$Sample


# Get total arg counts and add to meta_data frame
res_count_raw <- apply(arg_counts,2,sum)
new_meta_data <- merge(new_meta_data, as.data.frame(res_count_raw), by=0, all=TRUE)
rownames(new_meta_data) <- new_meta_data$Sample
new_meta_data <- new_meta_data %>% dplyr::select(-Row.names)

# Normalize total arg counts
info_frame <- new_meta_data %>% mutate(res_count_frag_norm=res_count/Mfragments)
info_frame <- info_frame %>% mutate(res_count_bac_norm=res_count/bacterial_16s)
info_frame <- info_frame %>% mutate(intI1_count_bac_norm=intI1/bacterial_16s)

# Merge the metagenomic data and the isolate/clinical data
info_frame <- merge(info_frame,isolate_merged,by="Country")
```

## Create phyloseq objects with the ARGs

```{r}
# Create phyloseq objects
frag_norm_arg <- length_normalized_args
arg_taxa <- as.matrix(arg_lineage)
rownames(arg_taxa) <- arg_lineage$ARG
colnames(arg_taxa)


my_meta_data <- sample_data(info_frame)
sample_names(my_meta_data) <- my_meta_data$Sample

# Create one phyloseq normalized by length and total million fragments
ps_args_frag_norm <- phyloseq(otu_table(frag_norm_arg, taxa_are_rows=TRUE), tax_table(arg_taxa),my_meta_data)
otu_table(ps_args_frag_norm) <- otu_table(sweep(otu_table(ps_args_frag_norm), 2, sample_data(ps_args_frag_norm)$Mfragments, FUN="/"),taxa_are_row=TRUE)
# Convert to FPKM
otu_table(ps_args_frag_norm) <- otu_table(ps_args_frag_norm)*1000


# Create one phyloseq normalized by length and bacterial 16s
ps_args_bac_norm <- phyloseq(otu_table(frag_norm_arg, taxa_are_rows=TRUE), tax_table(arg_taxa),my_meta_data)
otu_table(ps_args_bac_norm) <- otu_table(sweep(otu_table(ps_args_bac_norm), 2, sample_data(ps_args_bac_norm)$bacterial_16s, FUN="/"),taxa_are_row=TRUE)


# Create one with just the counts 
ps_args_count <- phyloseq(otu_table(arg_counts, taxa_are_rows=TRUE), tax_table(arg_taxa),my_meta_data)
```

# Remove variables

```{r}
semar_info_frame <- info_frame
variables_to_be_saved <- c("ps_args_count","ps_args_bac_norm","ps_args_frag_norm","semar_info_frame")
rm(list=setdiff(ls(), variables_to_be_saved))
```

# Load qPCR data

```{r}
# Define file paths
qpcr_count_file <- 'data/qpcr_data.txt'


# Read data from files
arg_data <- read_delim(qpcr_count_file, '\t')   # Reads qpcr count file with tab delimiter

# Select columns from arg_data
arg_data <- arg_data %>% dplyr::select(-Rya)    # Removes the 'Rya' column from arg_data

# Create a separate count for intI1 and remove it from arg_data
qpcr_intI_count <- arg_data %>% filter(group == "Integrons")
arg_data <- arg_data %>% filter(!group == "Integrons")

# Define lineage and assay information
lineage_to_use <- c("group", "gene", "assay")     # Specify the lineage columns to use
arg_lineage <- arg_data %>% dplyr::select(all_of(lineage_to_use))    # Select lineage columns from arg_data
#arg_names <- arg_data %>% dplyr::select(assay)    # Select the 'assay' column from arg_data

# Set row names for arg_data and arg_numbers_only
rownames(arg_data) <- arg_data$assay                 
arg_numbers_only <- arg_data %>% dplyr::select(-c("group", "gene", "assay"))   # Remove lineage columns from arg_data
rownames(arg_numbers_only) <- arg_data$assay      
```
Get some summaries such as total ARG count and fix with the metadata file

```{r}
# Get total arg counts normalized and add to meta_data frame
qpcr_count_norm <- apply(arg_numbers_only,2,sum)

#row.names(meta_data) <- meta_data$Country
res_frame <- data.frame(qpcr_count_norm)
res_frame <- res_frame %>% mutate(Country = rownames(res_frame))
semar_info_frame <- merge(semar_info_frame, res_frame, by="Country")

# Get intI1 counts normalized and add to meta_data frame
# get only counts
intI_counts <- qpcr_intI_count %>% dplyr::select(-all_of(lineage_to_use ))
qpcr_intI_norm <- apply(intI_counts,2,sum)

#row.names(meta_data) <- meta_data$Country
res_frame <- data.frame(qpcr_intI_norm)
res_frame <- res_frame %>% mutate(Country = rownames(res_frame))
semar_info_frame <- merge(semar_info_frame, res_frame, by="Country")

```

Create phyloseq ibhect for qPCR

```{r}
# Create phyloseq objects
frag_norm_arg <- arg_numbers_only
arg_taxa <- as.matrix(arg_lineage)
rownames(arg_taxa) <- arg_lineage$assay


# Change sample names to country
my_meta_data <- sample_data(semar_info_frame)
sample_names(my_meta_data) <- my_meta_data$Country

# Create one phyloseq normalized by length and total million fragments
ps_qpcr_norm <- phyloseq(otu_table(frag_norm_arg, taxa_are_rows=TRUE), tax_table(arg_taxa),my_meta_data)
```


# Remove and save variables

```{r}
variables_to_be_saved <- c("ps_args_count","ps_args_bac_norm","ps_args_frag_norm","semar_info_frame","ps_qpcr_norm","qpcr_intI_count")
rm(list=setdiff(ls(), variables_to_be_saved))
```